$3$-D Weak Lensing Mass Map Reconstruction: Reducing the line of sight smearing

### 2020-12-06 doc/paper_ms_method_HSCY1/sparse_NFW_results.tex
revising the paper

\begin{figure*}
 \centering
 \includegraphics[width=1.0\textwidth]{peak_histograms_NFW.pdf}
 \caption{The histograms of detected peak values (normalized) from all of the
     simulations.  The solid red steps are the results from reconstructions
     with the NFW dictionary penalized with different regularization
     parameters: $\lambda=3.5,4.0,5.0$. The dashed blue steps are the
     corresponding results of the reconstructions from $1000$ realizations of
     pure noise fields. The gray lines are the best-fit Gaussian distributions
     to the noises' peak histograms.
    }\label{fig_peakHist}
\end{figure*}

In this subsection, we test the performance of our algorithm with the default
setup that models the matter density field with multi-scale NFW atoms. The
dictionary is constructed with three frames of different NFW comoving scale
radius: $0.12~h^{-1}$ Mpc, $0.24~h^{-1}$ Mpc, and $0.36~h^{-1}$ Mpc.  The
truncation radii (concentration) are set to four times the comoving scale radii
for the atoms in the dictionary. We note that each frame of our dictionary
fixes the scale radius in the comoving space; therefore, the NFW atoms have
different angular radii in different lens redshift bins.

We test the algorithm with different regularization parameters for the
preliminary lasso estimation, which are $3.5$, $4.0$, and $5.0$. The
corresponding regularization parameters for the adaptive lasso are set to
$\lambda_{\rm{als}}=\lambda_{\rm{ls}}^{\tau+1}$.  Here, we note that both the
preliminary lasso estimation and the final adaptive lasso estimation select the
pixels with the signal-to-noise ratios (SNRs) greater than $\lambda_{\rm{ls}}$
in each gradient descent iteration. While the final adaptive lasso estimation
further enhances the growth of the pixels with preliminary estimations greater
than $\lambda_{\rm{ls}}$.

@@

The left panels of Figure \ref{fig_NFWvsPM} demonstrates the $3$-D density maps
reconstructed with different penalization parameters for a halo with
$M_{200}=10^{15.02} ~h^{-1}M_{\odot}$ at redshift $0.164$.  As we can see from
the reconstructed maps, the adaptive lasso algorithm sets most of the
reconstructed pixels to zero and only keeps the pixels with large amplitudes.
Therefore, it is straightforward to identify the peaks on the reconstructed
sparse density map.  Figure \ref{fig_peakHist} shows the histograms for the
detected peaks with different penalization parameters.

We simulate $1000$ realizations of pure noise fields and perform the
same reconstructions on these noise fields to study the noise properties. Figure
\ref{fig_peakHist} shows the histograms of peaks detected from the
pure noise fields. , we also find that the number of false
peaks significantly decreases as the penalization parameter increases.

@@

The stacked $2$-D histogram, from all of the simulations, for the offsets of
the detected peak positions from the input halos' positions is shown in the
left panel of Figure \ref{fig_detoffsets}.

For each simulation, we find the positive peak closest to the input position
(in the pixel unit). If the closest peak lay inside the region denoted with
the dashed box in the left panel of Figure \ref{fig_detoffsets}, we take it as
a true peak detection of the input halo. Other detected peaks, which include both
positive and negative peaks, are taken as false peaks. The right
panel of Figure \ref{fig_detoffsets} shows the average redshift of true
detections for each halo. The estimated redshifts are lower than the true
redshifts by about $0.03$ for halos in the low-redshift range ($z\leq 0.4$).
For halos at $0.4<z\leq 0.85$, the relative redshift bias is below $0.5\%$.

\begin{figure*}
 \centering
 \includegraphics[width=1.0\textwidth]{detfalse_threshold_NFW_lbd35.pdf}
 \includegraphics[width=1.0\textwidth]{detfalse_threshold_NFW_lbd40.pdf}
 \includegraphics[width=1.0\textwidth]{detfalse_threshold_NFW_lbd50.pdf}
 \caption{The upper panels show the halo detection rate and the number of false
     peaks per square degree for mass map reconstructed with the NFW atoms. The
     penalization parameter is set to $\lambda=3.5$ for the upper panels.  The
     low panels show the results for $\lambda=5.0$.
        } \label{fig_detFalsRateNFW}
\end{figure*}

Figure \ref{fig_detFalsRateNFW} shows the detection rate (left panels) and the
number of false peaks per square degree (right panel) for each simulated halo.
The figure include results of different penalization parameters: $\lambda=3.5$
(upper panel) and $\lambda=5.0$ (lower panel).  We find that with the increase
of the penalization parameter, the detection rate decreases. On the other hand,
the number of false peaks also decreases.

>>>>

\begin{figure*}
\centering
\subfigure[NFW: $\lambda=3.5$]{
        \includezraphics{delta-1-7-pz-wn-NFW-falsepeakproblem.pdf}
    }
\subfigure[NFW: $\lambda=5.0$]{\includezraphics{
        delta-1-7-pz-wn-NFW_3-falsepeakproblem.pdf}
    }
\caption{The density maps reconstructed from the mock galaxy shape catalog with
    the NFW dictionary. The penalization parameters are $\lambda=3.5$ (left)
    and $\lambda=5.0$ (right).  The input halo mass is $M_{200}=10^{15.02}
    ~h^{-1}M_{\odot}$, and its redshift is $z=0.164$.  The vertical direction
    is the line of sight direction. The boxes' lower boundaries and upper
    boundaries of correspond to $z=0.01$ and $z=0.85$, respectively.
    } \label{fig_NFW3D}
\end{figure*}

In this subsection, we test the performance of our algorithm with the default
setup that models the matter density field with multi-scale NFW atoms. The
dictionary is constructed with three frames of different NFW scale radii in the
comoving coordinate: $0.12~h^{-1}$ Mpc, $0.24~h^{-1}$ Mpc, and $0.36~h^{-1}$
Mpc.  The truncation radii are set to four times the comoving scale radii for
the atoms in the dictionary (concentration equals four) . We note that each
frame of our dictionary fixes the scale radius in the comoving space;
therefore, the NFW atoms have different angular radii in different lens
redshift bins.

We test the algorithm with different regularization parameters ($\lambda$) for
the preliminary lasso estimation, which are $3.5$, $4.0$, and $5.0$. The
corresponding regularization parameters for the final adaptive lasso estimations
are set to $\lambda_{\rm{ada}}=\lambda^{\tau+1}$.  Here, we note that both the
preliminary lasso estimation and the final adaptive lasso estimation select the
pixels with the signal-to-noise ratios (SNRs) greater than $\lambda$ in each
gradient descent iteration. While the final adaptive lasso estimation further
enhances the growth of the pixels with preliminary estimations greater than
$\lambda$.

@@

Figure \ref{fig_NFW3D} shows the $3$-D density maps reconstructed with
different penalization parameters for a halo with $M_{200}=10^{15.02}
~h^{-1}M_{\odot}$ at redshift $0.164$.  As we can see from the reconstructed
density maps, the adaptive lasso algorithm sets most of the reconstructed
pixels to zero and only keeps the pixels with large amplitudes.  We identify
the peaks on the reconstructed sparse density map.

\begin{figure*}
 \centering
 \includegraphics[width=1.0\textwidth]{peak_histograms_NFW.pdf}
 \caption{The number per square degree histograms of detected peak values from
     all of the simulations.  The solid red steps result from reconstructions
     with the NFW dictionary penalized with different regularization
     parameters: $\lambda=3.5,4.0,5.0$. The dashed blue steps are the
     corresponding results of the reconstructions from $1000$ realizations of
     pure noise fields. The gray lines are the best-fit Gaussian distributions
     to the noises' peak histograms.
    }\label{fig_peakHist}
\end{figure*}

Following \citet{WL-massMap-Glimpse2D-Lanusse2016}, we normalize detected peaks
in the $l$-th lens redshift plane to account for the peak amplitude difference
due to the difference in the norm of the lensing kernels for different redshift
bins:
\begin{equation}
\delta^{\rm{n}}_{\rm{peak}}(\vec{\theta},z_l)=\delta_{\rm{peak}}(\vec{\theta},z_l)/\mathcal{R}_{l}^{\frac{1}{2}},
\end{equation}
where the normalization matrix is defined as
\begin{equation}
\mathcal{R}_{l}=\sum_s K^2(z_l,z_s).
\end{equation}
The solid steps in Figure \ref{fig_peakHist} show the histograms for the
normalized peaks with different penalization parameters. Also, we simulate
$1000$ realizations of pure noise fields and perform the reconstructions on
these noise fields to study the noise properties. The dashed steps in Figure
\ref{fig_peakHist} show the histograms of normalized peaks detected from the
pure noise fields. The solid lines in Figure \ref{fig_peakHist} show the
best-fit Gaussian distributions to the noise peaks' histograms.

Figure \ref{fig_peakHist} tells that the densities of peaks (including both
true and false peaks) are suppressed as the penalization parameter $\lambda$
increases. Moreover, we find the standard deviation of noise peaks slightly
decreases as $\lambda$ increases. As a result, for a higher detection threshold
($\lambda=5.0$), we find a clearer peak number excess for mass maps
reconstructed from mock catalogs comparing with the noise peak histogram .

The $2$-D histogram, stacked from all of the simulations, for the offsets of
the detected peak positions from the input halos' positions is shown in the
left panel of Figure \ref{fig_detoffsets}. We see a clustering of peaks close
to the input halo's position on the stacked position histogram.  For each stamp
simulation, we find the positive peak closest to the input position (in the
pixel unit). If the closest peak lay inside the region denoted with the dashed
box in the left panel of Figure \ref{fig_detoffsets}, we take it as a true peak
detection of the input halo. Other identified peaks, which include both positive
and negative peaks, are taken as false detections.

@@

The right panel of Figure \ref{fig_detoffsets} shows the average redshift of
true detections for each halo. The estimated redshifts are lower than the true
redshifts by about $0.03$ for halos in the low-redshift range ($z\leq
0.4$).  For halos at $0.4<z\leq 0.85$, the relative redshift bias is below
$0.5\%$.

We select peaks with values greater than an ad-hoc threshold as candidates of
galaxy clusters following \citet{HSC1-massMap-cluster}.  The threshold is set
to a few times the standard deviation of the noise peaks. We use different
detection thresholds ($1.5\sigma$ and $3.0\sigma$) to detect galaxy clusters
from the mass maps reconstructed with $\lambda=3.5,4.0,5.0$. The left and
middle columns of Figure \ref{fig_detFalsRateNFW} show the detection rates for
halos in the (mass, redshift) planes with detection thresholds set to
$1.5\sigma$ and $3.0\sigma$ of the noise field, respectively. The right column
of Figure \ref{fig_detFalsRateNFW} shows the corresponding numbers of false
detections per square degree as functions of detection thresholds. The first,
second, and third rows of Figure \ref{fig_detFalsRateNFW} correspond to the
$lambda=3.5,4.0,5.0$, respectively.

Figure \ref{fig_detFalsRateNFW} tells that the false peak density is suppressed
as the detection threshold increases. Also, the detection rate of halo
significantly decreases. We decide to set the detection threshold to
$1.5\sigma$ and set the penalization parameter $\lambda$ to $5.0$ since such a
setup suppresses the false detection to $0.022$ while keeping a good halo
detection rate. In summary, The algorithm is able to detect halo with minimal
mass limits of $10^{14.0} M_{\odot}/h$, $10^{14.7} M_{\odot}/h$, $10^{15.0}
M_{\odot}/h$ for the low ($z<0.3$), median ($0.3\leq z< 0.6$) and high
($0.6\leq z< 0.85$) redshifts, respectively.

\begin{figure*}
 \centering
 \includegraphics[width=1.0\textwidth]{detfalse_threshold_NFW_lbd35.pdf}
 \includegraphics[width=1.0\textwidth]{detfalse_threshold_NFW_lbd40.pdf}
 \includegraphics[width=1.0\textwidth]{detfalse_threshold_NFW_lbd50.pdf}
 \caption{The detection rate and false peak density for different $\lambda$ and
     detection thresholds. The first, second, and third rows correspond to the
     results with $lambda=3.5,4.0,5.0$, respectively. The left and middle columns
     are the halo detection rates for detection thresholds equal $1.5\sigma$ and
     $3.0\sigma$, respectively. The right column is the density of false peaks as
     a function of detection threshold.
    } \label{fig_detFalsRateNFW}
\end{figure*}

### 2020-12-06 doc/paper_ms_method_HSCY1/sparse_PointMass_results.tex
revising the paper

In this subsection, we show the results for the setup, which substitutes the
default NFW dictionary with the point mass dictionary, to compare with the
default setup.  The regularization parameter for the preliminary lasso is set
to $3.5$ and $5.0$.  Inspired by \citet{structureAdaLasso-Pramanik2020}, which
propose to incorporate external group information into different adaptive lasso
penalization weights by setting the penalization weights for projection
coefficients in the same group to the average of the adaptive weights in this
group, we smooth the preliminary lasso estimation in each lens redshift bin
with a top-hat filter of comoving scale $0.25~h^{-1}$ Mpc. The smoothed
preliminary lasso estimation is denoted as $\hat{x}^{\rm{ls}}_{\rm{sm}}$, and
the penalization weights are set to $1/\abs{\hat{x}^{\rm{ls}}_{\rm{sm}}}^\tau$.
The regularization parameter for the adaptive lasso is set to
$\lambda_{\rm{als}}=\lambda_{\rm{ls}}$.

As we have done on the NFW dictionary's reconstruction, we smooth the
reconstructed density maps with the Gaussian kernel (scale radius equals
$1.\arcmin 5$) in each lens redshift plane.

Figure \ref{fig_detFalsRatePM} shows the detection rate (left panels) and the
number of false peaks per square degree (right panel) for each setup of halo
mass and redshift.
This figure includes results of different penalization parameters:
$\lambda=3.5$ (upper panel) and $\lambda=5.0$ (lower panel). The right panels
of Figure \ref{fig_NFWvsPM} show the $3$-D density maps reconstructed with
different penalization parameters for a halo with $M_{200}=10^{15.02}
~h^{-1}M_{\odot}$ at redshift $0.164$.
Comparing with the results of the NFW dictionary, we find that the number of
false peaks in the mass map reconstructed with the point mass dictionary is
much larger than the NFW dictionary.

As demonstrated in the right panels of Figure \ref{fig_NFWvsPM}, the mass
reconstructions with the point mass dictionary tend to assign masses to
several different redshift bins in the neighboring region of the halo center.
While the NFW dictionary produces consistent mass reconstructions.
Since the NFW atoms' profiles are more consistent with the mass profiles of the
input halos than the profile of a point mass, there are less false structures
reconstructed with the NFW dictionary.

>>>>

We substitute the default NFW dictionary with the point mass dictionary and
reconstruct the mass map from the mock galaxy shape catalog, to compare with
the default setup.

The regularization parameter ($\lambda$) for the preliminary lasso is set to
$3.5$ and $5.0$.  Inspired by \citet{structureAdaLasso-Pramanik2020}, which
propose to incorporate external group information into different adaptive lasso
penalization weights by setting the penalization weights for projection
coefficients in the same group to the average of the adaptive weights in this
group, we smooth the preliminary lasso estimation in each lens redshift bin
with a top-hat filter of comoving scale $0.25~h^{-1}$ Mpc. The smoothed
preliminary lasso estimation is denoted as $\hat{x}^{\rm{ls}}_{\rm{sm}}$, and
the penalization weights are set to
$\hat{w}=1/\abs{\hat{x}^{\rm{ls}}_{\rm{sm}}}^\tau$.
The regularization parameter for the adaptive lasso is set to
$\lambda_{\rm{ada}}=\lambda$.
As we have done on the NFW dictionary's reconstruction, we smooth the
reconstructed density maps with the Gaussian kernel (scale radius equals
$1.\arcmin 5$) in each lens redshift plane.

As demonstrated in Figure \ref{fig_PM3D}, the mass reconstructions with the
point mass dictionary tend to assign masses to several different redshift bins
in the neighboring region of the halo's center.  In contrast, as demonstrated in
Figure \ref{fig_NFW3D}, the NFW dictionary manages to perform consistent mass
reconstructions.  We think the problem of the point mass dictionary originates
from the fact that the profile of the point mass atom in the transverse plane
is much more compact than the profile of the input halos, especially at low
redshift.

\begin{figure*}
\centering
\subfigure[Point Mass: $\lambda=3.5$]{
        \includezraphics{delta-1-7-pz-wn-PM-falsepeakproblem.pdf}
        }
\subfigure[Point Mass: $\lambda=5.0$]{
        \includezraphics{delta-1-7-pz-wn-PM_3-falsepeakproblem.pdf}
        }
\caption{The density maps reconstructed from the mock galaxy shape catalog with
    the point mass dictionary. The penalization parameters are $\lambda=3.5$
    (left) and $\lambda=5.0$ (right).  The input halo mass is
    $M_{200}=10^{15.02} ~h^{-1}M_{\odot}$, and its redshift is $z=0.164$.  The
    vertical direction is the line of sight direction. The boxes' lower
    boundaries and upper boundaries of correspond to $z=0.01$ and $z=0.85$,
    respectively.
    } \label{fig_PM3D}
\end{figure*}
